---
title: "Let's Learn Dplyr!"
author: "Jessica McKay"
date: "10/8/2021"
output: word_document
---

## Chunk 1
First, it's important for us to set up our markdown file. Let's do that!
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chunk 2
Now, let's download our data and get ready to learn about penguins!
```{r data and packages, include=FALSE}

#### Load in the data 
install.packages("palmerpenguins")
library(palmerpenguins)
path_to_file("penguins_raw.csv")
data <- penguins
 
### Load in the needed packages
install.packages("tidyverse")
library(tidyverse)
install.packages("dplyr")
library(dplyr)
install.packages("ggplot2")
library(ggplot2)
```

### Chunk 3
Now, that we have the dplyr package downloaded, we can start to use pipes (%>%). 
Let's check out that data! Run the line before to view the data. 

```{r view data, include=FALSE}
#### Load in the data 
data %>%
  view()
```

### Chunk 4
That's pretty cool data! Now, I'm curious... How does the data change when we filter for just the female penguins?
```{r explore the data 1}

data %>%
  filter(sex == "female") %>%
  view()
```

### Chunk 5
hmmm, interesting! Now, let's just look at a few variables of interest.
```{r explore the data 2}

data %>%
  filter(sex == "female") %>%
  select(island, flipper_length_mm, bill_length_mm, sex) %>%
  view()

```

### Chunk 6
It would be easier to read through the data, if we arrange the data in a way that makes it easier to read through... Let's use the arrange command to do so!

```{r explore the data 3}
new_data <-  data %>%
  filter(sex == "female") %>%
  select(island, flipper_length_mm, bill_length_mm, sex, year) %>%
  arrange(island) %>%
  na.omit()
```


### Chunk 8
Now, let's take a look! How does it look?
```{r explore the data 4}
new_data %>%
  view()
```


### Chunk 9
Next, let's take a look at average flipper length among the penguins in our data. Be sure to use the summary command here! It will enable you to make neat little summary tables. 
```{r explore the data 5}

### New Data
new_data %>%
 group_by(sex) %>%
  
  summarize(mean(flipper_length_mm)) %>%
  na.omit() 


### Original Data

data %>%
  group_by(sex) %>%
  summarize(mean(flipper_length_mm)) %>%
  na.omit()
```

### Chunk 10
That is some interesting data... I wonder how average flipper length changed across years among our sample of female penguins? 
Let's create a neat summary table!
```{r explore the data 6}

 table_new_data  <- new_data %>%
    mutate(new_year = case_when(year == "2007" ~ "'07", year == "2008" ~ "`08", year =="2009" ~ "`09")) %>%
    group_by(`Year` = new_year) %>%
    summarize(`Average Flipper Length` = mean(flipper_length_mm)) 

```


### Chunk 11
Let's take a look. What trends do you see?
```{r explore the data 7}
 table_new_data  %>%
  print()
```

If time permits, let's see if we find similar trends for male penguins! Then, let's compare male and female penguins!




